### 2021-02-15 paper/natastron/Publication/methods.tex
references done fucking finally

For our asteroseismic ages, we used the ages obtained by \texttt{BASTA} \cite[BAyesian STellar Algorithm]{m_silvaaguirre+2015} in the `Kages' and LEGACY catalogues. These ages have been obtained by comparisons of measured oscillation properties to stellar models, accounting for an expanded range of metallicities. \texttt{BASTA} is thoroughly compared to four other seismic modelling techniques in \cite{silvaaguirre+2017}. While uncertainties found through \texttt{BASTA} are typically higher than for other techniques, only \texttt{BASTA} and \texttt{ASTFIT} \cite[Aarhus STellar Evolution Code]{m_christensen-dalsgaard2008} recover the radius, mass and age of the Sun, when applied to solar data. Although the uncertainties on \texttt{ASTFIT} ages are overall lower, they are not published for the `Kages' sample. In order to maintain an internally consistent stellar age sample, we used age results from \texttt{BASTA} for both the `Kages' and LEGACY samples.

@@

As described in the catalogue papers, for `Kages' stars atmospheric properties (\teff and \feh) were measured through high-resolution spectroscopy \cite{huber+2013a}. For LEGACY stars, atmospheric properties were similarly taken from one study \cite{buchhave+latham2015} for most stars in the catalogue, and complemented by other values from the literature for the remaining stars \cite[see Table 3]{m_silvaaguirre+2017}.

Other asteroseismic properties used in this study, such as those used as first guesses on the free parameters in our asteroseismic model, were taken from the `Kages' and LEGACY catalogues \cite{astropycollaboration+2013, m_astropycollaboration+2018, m_ginsburg+2019}. In cases where both catalogues contained the same target, we used the stellar parameters reported in LEGACY \cite{mckinney2010}. This is also the case for the masses and ages described above.

@@

In order to extract signatures of stellar rotation from the asteroseismic mode frequencies, we built a model that simultaneously treats the convective background, modes of oscillation, and white noise, while accounting for rotational splitting. The foundations of this approach follows best-practices in asteroseismology \cite{davies+2015}, fitting Harvey profiles for the background \cite{harvey1985} and expressing the frequencies of the modes using an asymptotic expression \cite{tassoul1980, m_vrard+2016}.

The first core improvement in this work is the use of hierarchical latent variable models \cite{hogg+2010, m_hall+2019}, which account for small-scale deviations in mode frequencies due to effects not explicitly accounted for in our model (such as acoustic glitches \cite{mazumdar+2014}). By improving the inference of mode frequencies in this way, we also improve the ability to resolve the rotational splitting. The second improvement comes in the form of priors on our parameters, and in particular the rotational inclination, which more accurately reflects the true distribution of angles than previous techniques \cite{chaplin+basu2017}. A step-by-step breakdown of the model and all priors can be found in \textbf{Supplementary Information}.\\

We separated our model fitting into two parts. First, we fit a model for the convective background and white noise only, to the region of the power spectra that did not contain modes of oscillation. This was done using \texttt{PyStan} \cite{vanhoey+2013, m_carpenter+2017}, for 10,000 iterations on each star. These \texttt{PyStan} runs were initiated with a random seed of 11, as were all other random processes in this work. Second, we fit our full model (including the convective background) to the region containing the modes of oscillation. The results of the first fit to the convective background were used as extremely informative priors on the background parameters in this second fit. This was done using \texttt{PyMC3} \cite{vanderwalt+2011,m_salvatier+2016, m_thetheanodevelopmentteam+2016} for 2500 iterations each on 4 chains.

@@

\section{Gyrochronology Methods}\label{s:gyro}
\subsection{Stellar Models}\label{ssec:models}
The braking models used in this work have several parameters that inform the rotational evolution. These are: a normalization factor to reproduce the solar rotation ($f_k$); a disk locking timescale ($T_{\rm disk}$) and period ($P_{\rm disk}$) which together regulate the stellar angular velocity during the pre-main sequence; the critical angular velocity that marks the transition from saturated (rapidly rotating) to unsaturated (slowly rotating) regimes ($\omega_{\rm crit}$), and the critical Rossby number, above which stars conserve angular momentum ($Ro_{\rm crit}$), mentioned above. $T_{\rm disk}$, $P_{\rm disk}$ and $\omega_{\rm crit}$ are calibrated to match the behavior in young open clusters, but have little impact on the rotational evolution beyond $\approx 1\, \rm{Gyr}$ in solar-mass stars. Both $f_k$ and $Ro_{\rm crit}$ affect the late-time evolution. Both models adopt $\omega_{\rm crit} = 3.4  \times 10^{-5}\, \rm{s}^{-1}$, $P_{\rm disk} = 8.1$, $T_{\rm disk} = 0.28$ and $f_k = 6.6$. In the weakened magnetic braking model, $Ro_{\rm crit} = 1.97$. \rtwo{This Rossby number is based on model comparisons to catalogues of spot rotation rates \cite{vansaders+2019}. Its exact value is not critical in our analysis, as any similar value would create a model population significantly different from the standard case. In the model, the Rossby number is defined as the rotation period divided by the convective turnover timescale, which in turn is defined as $\tau_{\rm cz} = H_p / v_{\rm conv}$, where $H_p$ is the pressure scale height at the base of the convection zone, and $v_{\rm conv}$ is the convetive velocity evaluated one pressure scale height above the base in a mixing length theory of convection. Solid body rotation is assumed \cite{nielsen+2015}.} For further details, see \cite{vansaders+pinsonneault2013} and \cite{vansaders+2016, m_vansaders+2019} for the standard and weakened magnetic braking cases respectively.

@@

This new distribution was used for the nearest-neighbour matching. Once drawn we dropped the binary companion and used the true TRILEGAL properties of those stars. For stars where the mass of the companion was $M < 0.4\, M_\odot$, binary contributions were ignored. Every binary was assumed to result in a blend, regardless of separation. This results in slightly more young stars than reality, because young, blended binary systems contaminate regions of the HR diagram where one expects to find old stars, and the number of blends is overestimated by assuming every binary system is a blend.

@@

In order to draw probabilistic inference about the models, we built a five-dimensional Kernel Density Estimate (KDE) of both model populations using the \texttt{statsmodels} package \cite{seabold+perktold2010}. We used a band-width (setting the resolution of the KDE) of $0.02\, M_\odot$ in mass, $10\, \rm K$ in \teff, $0.01\, \rm{dex}$ in $\ln(t)$, $0.01\, \rm{dex}$ in $\rm{[Fe/H]}$ and $0.01\, \rm{dex}$ in $\ln(P)$. Note that age and rotation were treated in log space, where the posterior estimates from asteroseismology more closely resemble normal distributions. This approach translates the population models to a probability distribution we can use in a Bayesian framework.

@@

Typically, this model would evaluate all stars in our sample against the stellar models simultaneously for a single posterior estimate of $Q_{\rm WMB}$. At 95 stars, in 5 parameter spaces, this totals 476 free parameters to marginalise over. This is not an issue for Hamiltonian Monte Carlo \cite[HMC]{m_betancourt+girolami2013}, however the use of KDE functions, over which a probabilistic gradient can not be measured, reduces HMCs effectiveness. Alternative Markov Chain Monte Carlo techniques \cite[MCMC]{m_foreman-mackey+2013} can more efficiently sample the KDE functions, but can not treat the large number of hierarchical parameters. To overcome this, we fit our model to each star to obtain an independent individual posterior distribution for $Q_{\rm WMB}$, and multiplied these post hoc to obtain a combined posterior. This comes with the benefit of easily allowing us to calculate the combined posterior for different stellar classifications, at the expense of the ability to marginalise for a single value of $Q_{\rm WMB}$ directly.\\

@@

We fit our model Equation \ref{eq:modelll} using \texttt{emcee} \cite{foreman-mackey+2013, m_foreman-mackey2016}, using 32 walkers for a total of 7500 samples per walker, of which the first 2500 were discarded as a burn-in.

@@

This work made use of publicly available data. \kepler power spectral densities were obtained from the \href{http://kasoc.phys.au.dk/}{KASOC} webpages for the majority of stars, and from the \href{https://archive.
		stsci.edu/prepds/kepseismic/}{MAST} for 16 Cyg A and B. This work used asteroseismic data from Silva Aguirre et al. (2015, 2017), Davies et al. (2016) and Lund et al. (2017) \cite{silvaaguirre+2015, m_silvaaguirre+2017,m_davies+2016, m_lund+2017}. Parameter distributions of the \kepler field used to alter our stellar population models were taken from Berger et al. (2020) \cite{berger+2020}.\\

@@

\input{bblmethods.tex}
%\bibliographystyle{naturemag}
%\bibliography{library.bib} % if your bibtex file is called example.bib

%\end{document}

>>>>

For our asteroseismic ages, we used the ages obtained by \texttt{BASTA} \cite[BAyesian STellar Algorithm]{silvaaguirre+2015} in the `Kages' and LEGACY catalogues. These ages have been obtained by comparisons of measured oscillation properties to stellar models, accounting for an expanded range of metallicities. \texttt{BASTA} is thoroughly compared to four other seismic modelling techniques in \cite{silvaaguirre+2017}. While uncertainties found through \texttt{BASTA} are typically higher than for other techniques, only \texttt{BASTA} and \texttt{ASTFIT} \cite[Aarhus STellar Evolution Code]{christensen-dalsgaard2008} recover the radius, mass and age of the Sun, when applied to solar data. Although the uncertainties on \texttt{ASTFIT} ages are overall lower, they are not published for the `Kages' sample. In order to maintain an internally consistent stellar age sample, we used age results from \texttt{BASTA} for both the `Kages' and LEGACY samples.

@@

As described in the catalogue papers, for `Kages' stars atmospheric properties (\teff and \feh) were measured through high-resolution spectroscopy \cite{huber+2013a}. For LEGACY stars, atmospheric properties were similarly taken from one study \cite{buchhave+latham2015} for most stars in the catalogue, and complemented by other values from the literature for the remaining stars \cite[see Table 3]{silvaaguirre+2017}.

Other asteroseismic properties used in this study, such as those used as first guesses on the free parameters in our asteroseismic model, were taken from the `Kages' and LEGACY catalogues \cite{astropycollaboration+2013, astropycollaboration+2018, ginsburg+2019}. In cases where both catalogues contained the same target, we used the stellar parameters reported in LEGACY \cite{mckinney2010}. This is also the case for the masses and ages described above.

@@

In order to extract signatures of stellar rotation from the asteroseismic mode frequencies, we built a model that simultaneously treats the convective background, modes of oscillation, and white noise, while accounting for rotational splitting. The foundations of this approach follows best-practices in asteroseismology \cite{davies+2015}, fitting Harvey profiles for the background \cite{harvey1985} and expressing the frequencies of the modes using an asymptotic expression \cite{tassoul1980, vrard+2016}.

The first core improvement in this work is the use of hierarchical latent variable models \cite{hogg+2010, hall+2019}, which account for small-scale deviations in mode frequencies due to effects not explicitly accounted for in our model (such as acoustic glitches \cite{mazumdar+2014}). By improving the inference of mode frequencies in this way, we also improve the ability to resolve the rotational splitting. The second improvement comes in the form of priors on our parameters, and in particular the rotational inclination, which more accurately reflects the true distribution of angles than previous techniques \cite{chaplin+basu2017}. A step-by-step breakdown of the model and all priors can be found in \textbf{Supplementary Information}.\\

We separated our model fitting into two parts. First, we fit a model for the convective background and white noise only, to the region of the power spectra that did not contain modes of oscillation. This was done using \texttt{PyStan} \cite{vanhoey+2013, carpenter+2017}, for 10,000 iterations on each star. These \texttt{PyStan} runs were initiated with a random seed of 11, as were all other random processes in this work. Second, we fit our full model (including the convective background) to the region containing the modes of oscillation. The results of the first fit to the convective background were used as extremely informative priors on the background parameters in this second fit. This was done using \texttt{PyMC3} \cite{vanderwalt+2011, salvatier+2016, thetheanodevelopmentteam+2016} for 2500 iterations each on 4 chains.

@@

\section{Gyrochronology Methods}\label{s:gyro}
\subsection{Stellar Models}\label{ssec:models}
The braking models used in this work have several parameters that inform the rotational evolution. These are: a normalization factor to reproduce the solar rotation ($f_k$); a disk locking timescale ($T_{\rm disk}$) and period ($P_{\rm disk}$) which together regulate the stellar angular velocity during the pre-main sequence; the critical angular velocity that marks the transition from saturated (rapidly rotating) to unsaturated (slowly rotating) regimes ($\omega_{\rm crit}$), and the critical Rossby number, above which stars conserve angular momentum ($Ro_{\rm crit}$), mentioned above. $T_{\rm disk}$, $P_{\rm disk}$ and $\omega_{\rm crit}$ are calibrated to match the behavior in young open clusters, but have little impact on the rotational evolution beyond $\approx 1\, \rm{Gyr}$ in solar-mass stars. Both $f_k$ and $Ro_{\rm crit}$ affect the late-time evolution. Both models adopt $\omega_{\rm crit} = 3.4  \times 10^{-5}\, \rm{s}^{-1}$, $P_{\rm disk} = 8.1$, $T_{\rm disk} = 0.28$ and $f_k = 6.6$. In the weakened magnetic braking model, $Ro_{\rm crit} = 1.97$. \rtwo{This Rossby number is based on model comparisons to catalogues of spot rotation rates \cite{vansaders+2019}. Its exact value is not critical in our analysis, as any similar value would create a model population significantly different from the standard case. In the model, the Rossby number is defined as the rotation period divided by the convective turnover timescale, which in turn is defined as $\tau_{\rm cz} = H_p / v_{\rm conv}$, where $H_p$ is the pressure scale height at the base of the convection zone, and $v_{\rm conv}$ is the convetive velocity evaluated one pressure scale height above the base in a mixing length theory of convection. Solid body rotation is assumed \cite{nielsen+2015}.} For further details, see \cite{vansaders+pinsonneault2013} and \cite{vansaders+2016, vansaders+2019} for the standard and weakened magnetic braking cases respectively.

@@

This new distribution was used for the nearest-neighbour matching. Once drawn we dropped the binary companion and used the true TRILEGAL properties of those stars. For stars where the mass of the companion was $M < 0.4\, \odot$, binary contributions were ignored. Every binary was assumed to result in a blend, regardless of separation. This results in slightly more young stars than reality, because young, blended binary systems contaminate regions of the HR diagram where one expects to find old stars, and the number of blends is overestimated by assuming every binary system is a blend.

@@

In order to draw probabilistic inference about the models, we built a five-dimensional Kernel Density Estimate (KDE) of both model populations using the \texttt{statsmodels} package \cite{seabold+perktold2010}. We used a band-width (setting the resolution of the KDE) of $0.02\, \odot$ in mass, $10\, \rm K$ in \teff, $0.01\, \rm{dex}$ in $\ln(t)$, $0.01\, \rm{dex}$ in $\rm{[Fe/H]}$ and $0.01\, \rm{dex}$ in $\ln(P)$. Note that age and rotation were treated in log space, where the posterior estimates from asteroseismology more closely resemble normal distributions. This approach translates the population models to a probability distribution we can use in a Bayesian framework.

@@

Typically, this model would evaluate all stars in our sample against the stellar models simultaneously for a single posterior estimate of $Q_{\rm WMB}$. At 95 stars, in 5 parameter spaces, this totals 476 free parameters to marginalise over. This is not an issue for Hamiltonian Monte Carlo \cite[HMC]{betancourt+girolami2013}, however the use of KDE functions, over which a probabilistic gradient can not be measured, reduces HMCs effectiveness. Alternative Markov Chain Monte Carlo techniques \cite[MCMC]{foreman-mackey+2013} can more efficiently sample the KDE functions, but can not treat the large number of hierarchical parameters. To overcome this, we fit our model to each star to obtain an independent individual posterior distribution for $Q_{\rm WMB}$, and multiplied these post hoc to obtain a combined posterior. This comes with the benefit of easily allowing us to calculate the combined posterior for different stellar classifications, at the expense of the ability to marginalise for a single value of $Q_{\rm WMB}$ directly.\\

@@

We fit our model Equation \ref{eq:modelll} using \texttt{emcee} \cite{foreman-mackey+2013, foreman-mackey2016}, using 32 walkers for a total of 7500 samples per walker, of which the first 2500 were discarded as a burn-in.

@@

This work made use of publicly available data. \kepler power spectral densities were obtained from the \href{http://kasoc.phys.au.dk/}{KASOC} webpages for the majority of stars, and from the \href{https://archive.
		stsci.edu/prepds/kepseismic/}{MAST} for 16 Cyg A and B. This work used asteroseismic data from Silva Aguirre et al. (2015, 2017), Davies et al. (2016) and Lund et al. (2017) \cite{silvaaguirre+2015, silvaaguirre+2017, davies+2016, lund+2017}. Parameter distributions of the \kepler field used to alter our stellar population models were taken from Berger et al. (2020) \cite{berger+2020}.\\

@@

\input{bblmethods.tex}
%\bibliographystyle{naturemag}
%\bibliography{library.bib} % if your bibtex file is called example.bib
%
%\end{document}
